---
title: "NFD MINOR UPDATE"
output: html_notebook
---

# PACKAGE DOC

- It must be mentioned that this is my first major project and it solves a serious problem.
- On a lighter side, I need to write a gaming program next to begin to relax my mind on just making cool fun stuff.

## Reading Files In

```{r}
usethis::edit_r_environ()
dat = Sys.getenv('first_dat')
dat2 = Sys.getenv('scnd_dat')
dat3 = Sys.getenv('third_dat')
lipo_dat <- read.table(dat)
lipo_dat2 <- read.table(dat2)
lipo_dat3 <- read.table(dat3)
```

## First Function (Clean Normal DATS)

```{r}
clean_evendat <- function(df){
  for (i in (4 * (1:(nrow(df)/4)))){
    k <- seq(4)
    skip_values = 8 * seq(40)
    if(i %in% skip_values) next
    df[c(k+i,i),] <- NA
  }
  df <- na.omit(df)
  return(df)
}

clean_evendat(lipo_dat)
```

## Second Function (Clean Abnormal DATS)

```{r}
comma_cleaner <- function(comma_df){
  cols_to_becleaned <- c(colnames(comma_df))
  comma_df[ , cols_to_becleaned] <- lapply(comma_df[ , cols_to_becleaned],  # Convert data
                                           function(x){ as.numeric(as.character(gsub(",", "", x))) })
  return(as.data.frame(comma_df))

}
```

## Third Function (Clean Abnormal DATS)

```{r}
clean_odddat <- function(df){
  library(dplyr)
  special_chars <- c('-,','-' )
  for (i in 1:nrow(df)){
    for (j in 1:ncol(df)){
      if(special_chars[1] %in% df[i,j] || special_chars[2] %in% df[i,j]){
        df[i,j] <- NA
      }
    }
  }
  nona_rows_df <- df

  comma_df <- nona_rows_df
  nocomma_df <- comma_cleaner(comma_df)
  nocomma_df <- nocomma_df[rowSums(is.na(nocomma_df)) != ncol(nocomma_df), ]

  return(nocomma_df)
}

wrangled_dat <- clean_odddat(lipo_dat)
```

## Fourth Function (Clean abnormal DATS)

```{r}
clean_odd_cc <- function(df){

  df <- comma_cleaner(df)
  df <- df[rowSums(is.na(df)) != ncol(df), ]

  return(df)

}

wrangled_dat <- clean_odddat(lipo_dat)
```

## Fifth Function (Resampling samples)

```{r}
resample_dat <- function(df, tnp, cycles){

  type_size <- c(1:tnp)
  k <- c(1:tnp)

  resulting_df <- data.frame()
  for (i in 1:(nrow(df)/tnp)){

    colnames(resulting_df) = NULL
    insert_row = df[k,]
    colnames(insert_row) = NULL

    resulting_df[i,type_size] <- rbind(insert_row, resulting_df, showWarnings = FALSE)

    increment = tnp
    k <- k + increment


  }
  return(resulting_df)
}
```

## Sixth Function (Resampling samples scaled)

```{r}
resample_dat_scale <- function(df, tnp, cycles){
  library(data.table)
  library(dplyr)

  col_list <- c()
  for(i in 1:ncol(df)){
    n <- "a"
    col_list <- c(col_list,assign(paste0(n, i), as.data.frame(df[,i])) )
  }

  j_vect <- c()
  for(j in col_list){
    j <- as.data.frame(j)
    j_resampled <- resample_dat(j, tnp = tnp, cycles = cycles)
    j_dfs <- as.data.frame(j_resampled)
    j_vect <- c(j_vect, j_dfs)
  }

  big_data = do.call(rbind, j_vect)
  big_data = as.data.frame(big_data)
  big_data_t = transpose(l=big_data)

  big_data_t <- big_data_t %>% select_if(~ !any(is.na(.)))
  #big_data_t <- big_data_t[ , colSums(is.na(big_data_t))==0]

  return(big_data_t)
}

test_scale <- resample_dat_scale(test, tnp = 3, cycles = 40)

resample_dat_alt_scale <- function(df, tnp, cycles, samples_per_tnp=NULL){
  
  type_size <- c(1:ncol(df)) #k is now nseq(same kinda thing)
  
  sample_df <- data.frame()
  final_df <- matrix(ncol = cycles)
  
  for(i in 1:tnp){
    nseq <- c(i:ncol(df))
    nseq <- nseq
    
    for (j in 1:(nrow(df)/tnp)){
      
      insert_row = df[nseq,]
      sample_df[j,type_size] <- rbind(insert_row, sample_df)
      increment_n = tnp
      nseq <- nseq + increment_n
      
    }
    final_df <- cbind(final_df, sample_df)
    final_df <- final_df[ , colSums(is.na(final_df))==0]
    colnames(final_df) <- NULL
    rownames(final_df) <- c(1:cycles)
    final_df <- as.data.frame(final_df)
    
  }
  #return(sample_df)
  return(final_df)
}
resamp_demo <- resample_dat_alt_scale(test,3,40)
```

## Sixth Function (Naming attributes)

```{r}
dat_col_names_prime <- function(df, rows_used = NULL, cols_used= NULL, user_specific_labels = NULL, read_direction = NULL){

  if(is.null(rows_used)){
    warning('user must enter rows_used which is a character vector with length == tnp')
  }

  col_names <- c()
  col_names_sort <- c()
  if(!is.null(cols_used)){
    normal_sequence = c(min(cols_used):max(cols_used))
  } else {
    normal_sequence = NULL
  }

  if(!is.null(user_specific_labels)){
    return(user_specific_labels)

  } else if(is.null(cols_used)){
    for(i in 1:ncol(df)){
      col_names <- c(col_names, paste0(rows_used,i))
    }
    if(is.null(read_direction) || read_direction == 'vertical'){
      message('check data frame for NA column or last sample column names with unmatched or isolated samples')
      return(col_names[1:ncol(df)])
    } else if(read_direction == 'horizontal'){
      for(i in 1:(ncol(df)/length(rows_used)) ){
        col_names_sort <- c(col_names_sort, paste0(rows_used,i))
      }
      col_names_sort <- stringr::str_sort(col_names_sort, decreasing = F, na_last = T, locale = 'en', numeric = T)
      message('check data frame for NA column names or last sample column with unmatched or isolated samples')
      return(col_names_sort[1:ncol(df)])
    }

  } else if(!is.null(rows_used) && !is.null(cols_used) && ncol(df) == length(cols_used)*length(rows_used) && length(cols_used) <= length(normal_sequence) ){

    for(i in cols_used){
      col_names <- c(col_names, paste0(rows_used,i))
    }
    if(is.null(read_direction) || read_direction == 'vertical'){
      return(col_names[1:ncol(df)])
    } else if(read_direction == 'horizontal'){
      col_names_sort <- stringr::str_sort(col_names, decreasing = F, na_last = T, locale = 'en', numeric = T)
      return(col_names_sort[1:ncol(df)])
    }
  } else if(is.null(user_specific_labels) || ncol(df) < length(cols_used)*length(rows_used) && length(cols_used) < length(normal_sequence) ){
    if(ncol(df) > length(cols_used)*length(rows_used)){
      message('number of columns exceeds users estimate; try leaving the cols_used blank')
      for(i in cols_used){
        col_names <- c(col_names, paste0(rows_used,i))
      }
      if(is.null(user_specific_labels) && is.null(read_direction) || read_direction == 'vertical'){
        print(col_names[1:ncol(df)] )
        print('From the printed above list enter the columns used; must match the sample positions on the plate;')
        choose_cols_used=scan(what=character(), n=ncol(df))
        print(choose_cols_used)
        return(as.vector(choose_cols_used))
      } else if(is.null(user_specific_labels) && !is.null(read_direction) && read_direction == 'horizontal'){
        col_names_sort <- stringr::str_sort(col_names, decreasing = F, na_last = T, locale = 'en', numeric = T)
        print(col_names_sort[1:ncol(df)] )
        print('From the printed above list enter the columns used; must match the sample positions on the plate;')
        choose_cols_used=scan(what=character(), n=ncol(df))
        print(choose_cols_used)
      }else{
        return(user_specific_labels)
      }

    }else {
      message('number of columns is lower than the users estimate; select the columns used from the list below')
      for(i in cols_used){
        col_names <- c(col_names, paste0(rows_used,i))
      }
      if(is.null(user_specific_labels) && is.null(read_direction) || read_direction == 'vertical'){
        print(col_names[1:(length(cols_used)*length(rows_used))])
        print('From the printed above list enter the columns used; must match the sample positions on the plate;')
        choose_cols_used=scan(what=character(), n=ncol(df))
        print(choose_cols_used)
        return(as.vector(choose_cols_used))
      } else if(is.null(user_specific_labels) && !is.null(read_direction) && read_direction == 'horizontal'){
        col_names_sort <- stringr::str_sort(col_names, decreasing = F, na_last = T, locale = 'en', numeric = T)
        print(col_names_sort[1:(length(cols_used)*length(rows_used))])
        print('From the printed above list enter the columns used; must match the sample positions on the plate;')
        choose_cols_used=scan(what=character(), n=ncol(df))
        print(choose_cols_used)
      }else{
        return(user_specific_labels)
      }
    }

  }
}
```


## Seventh Function (Fluor threshold check)

```{r}
check_max_fluor <- function(clean_df){
  library(emojifont)
  load.emojifont(font = "EmojiOne.ttf")
  for(i in 1:nrow(clean_df)){
    for(j in 1:ncol(clean_df)){
      if( (clean_df[i,j] >= 2^15) || (clean_df[i,j] <= 2^11 ) && !is.null(clean_df[i,j]) ){
        message(c(emoji('pig'), emoji('camel')))
        message("YIKES, one or more of your fluorescence values is greater than 2^16(65536) or less than 2^11(2048), watchout for very high fluorescence or very low in your next experimental design")
        message(paste('these values are either too high or low and can lead to NOISE','column:', j , 'row:', i))
      }
    }
  }
}
check_max_fluor(trial1_scale)
```

## Eight Function (Normalizer)

```{r}
min_max_norm <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}
```

## Ninth Function (Unique identifier cycle No)

```{r}
unique_identifier <- function(df){
  for(i in 1:nrow(df)){
    x <- 0
    x <- x + i
    df$unique_id[i] <- x
  }
  return(df)
}
```

## Tenth Function (Normfluodat)

```{r}
n = c('A','B','C')

normfluodat <- function(dat, tnp, cycles, rows_used = NULL, cols_used= NULL, user_specific_labels = NULL, read_direction = NULL){

  library(data.table)

  df <- read.table(dat) #dat becomes df
  df <- clean_odddat(df)
  df <- comma_cleaner(df)
  df <- as.data.frame(df)

  col_list <- c()
  for(i in 1:ncol(df)){
    n <- "a"
    col_list <- c(col_list,assign(paste0(n, i), as.data.frame(df[,i])) )
  }

  j_vect <- c()
  for(j in col_list){
    j <- as.data.frame(j)
    j_resampled <- resample_dat(j, tnp = tnp, cycles = cycles)
    j_dfs <- as.data.frame(j_resampled)
    j_vect <- c(j_vect, j_dfs)
  }

  cleaned_dat = do.call(rbind, j_vect)
  cleaned_dat = as.data.frame(cleaned_dat)
  cleaned_dat_t = transpose(l=cleaned_dat)

  #normalize
  cleaned_dat_t <- as.data.frame(lapply(cleaned_dat_t[1:ncol(cleaned_dat_t)], min_max_norm))

  #name the columns
  ru = rows_used
  cu = cols_used
  usl = user_specific_labels
  dr = read_direction
  sample_col_names <- dat_col_names_prime(cleaned_dat_t, ru, cu, usl, dr)
  colnames(cleaned_dat_t) <- sample_col_names

  #add unique_id
  cleaned_dat_t <-unique_identifier(cleaned_dat_t)

  return(cleaned_dat_t)
}
#a look at the logic of the failing normfluodat. NA omit is ran too early presenting the issue of truncating samples earlier than expected when the user does not load wells in sequence. Given this really happens when using the bad clean_odddat function.

normfluodat <- function(dat, tnp, cycles, rows_used = NULL, cols_used= NULL, user_specific_labels = NULL, read_direction = NULL){

  library(data.table)

  df <- read.table(dat) #dat becomes df
  df <- clean_odd_cc(df)

  col_list <- c()
  for(i in 1:ncol(df)){
    n <- "a"
    col_list <- c(col_list,assign(paste0(n, i), as.data.frame(df[,i])) )
  }

  j_vect <- c()
  for(j in col_list){
    j <- as.data.frame(j)
    j_resampled <- resample_dat(j, tnp = tnp, cycles = cycles)
    j_dfs <- as.data.frame(j_resampled)
    j_vect <- c(j_vect, j_dfs)
  }

  cleaned_dat = do.call(rbind, j_vect)
  cleaned_dat = as.data.frame(cleaned_dat)
  cleaned_dat_t = transpose(l=cleaned_dat)
  cleaned_dat_t <- cleaned_dat_t %>% select_if(~ !any(is.na(.)))
  check_max_fluor(cleaned_dat_t)

  #normalize
  cleaned_dat_t <- as.data.frame(lapply(cleaned_dat_t[1:ncol(cleaned_dat_t)], min_max_norm))

  #name the columns
  ru = rows_used
  cu = cols_used
  usl = user_specific_labels
  rd = read_direction
  sample_col_names <- dat_col_names_prime(cleaned_dat_t, ru, cu, usl, rd)
  colnames(cleaned_dat_t) <- sample_col_names

  #add unique_id
  cleaned_dat_t <-unique_identifier(cleaned_dat_t)

  return(cleaned_dat_t)
}

```




